Characterizing dynamical transitions by statistical complexity measures based on ordinal pattern transition networks

### 2020-09-29 manuscript_Huang.tex
Updates from Overleaf

\author{Reik V. Donner}
        \affiliation{Department of Water, Environment, Construction and Safety, Magdeburg--Stendal University of Applied Sciences, Breitscheidstra{\ss}e 2, 39114 Magdeburg, Germany}
        \affiliation{Potsdam Institute for Climate Impact Research (PIK) -- Member of the Leibniz Society, Telegrafenberg A31, 14473 Potsdam, Germany}

@@

\begin{abstract}
In the recent decade, complex network approaches are emerging as complementary novel ways to perform nonlinear time series analysis, in particular showing many insights that are hidden in the traditional methods. In this work, we focus on one of particular method of characterizing ordinal pattern transition network for time series data. More specifically, we generalize some traditional estimators of statistical complexity measures (SCM), which explicitly disclosing inhomogeneous frequencies of ordinal pattern transitions in the time series networks. To show the usefulness of the estimators, we use SCM to characterize different dynamical transitions, as illustrated by examples from the numerical model of  Logistic map and real time series of fluid experiments and ECG data. Both results of numerical and experimental data demonstrate that the transition frequencies between different ordinal patterns improve the SCM estimations, which are especially important for finite length of time series analysis, hence providing substantial information to the existing complex network approaches for nonlinear time series analysis. 
\end{abstract}

@@

@@

\section{Acknowledgements}
Parts of this work have been financially supported by the National Natural Science Foundation of China (Grant No. 11872182 and 11835003). Y. Z. further acknowledges the support by Shanghai Municipal Science and Technology Major Project (No.2018SHZDZX01) and ZJ Lab.

>>>>

\author{Reik V. Donner}
        \affiliation{Department of Water, Environment, Construction and Safety, Magdeburg--Stendal University of Applied Sciences, Breitscheidstra{\ss}e 2, 39114 Magdeburg, Germany}
        \affiliation{Potsdam Institute for Climate Impact Research (PIK) -- Member of the Leibniz Association, Telegrafenberg A31, 14473 Potsdam, Germany}

@@

\begin{abstract}
In the recent decade, complex network approaches have been emerging as novel and complementary concepts of nonlinear time series analysis that are able to unveil many features that are hidden to more traditional analysis methods. In this work, we focus on one particular approach: the application of ordinal pattern transition network for characterizing time series data. More specifically, we generalize some traditional estimators of statistical complexity measures (SCM), which explicitly disclose heterogeneous frequencies of ordinal pattern transitions in the time series networks. To demonstrate the usefulness of the estimators, we employ SCM to characterize different dynamical transitions in the paradigmatic model system of the logistic map as well as real-world time series of fluid experiments and ECG data. The obtained results for both, artificial and experimental data demonstrate that the consideration of transition frequencies between different ordinal patterns improves the estimation of SCM, which is especially important for the analysis of time series with finite length. 
\end{abstract}

@@

\textcolor{red}{If this paper is dedicated for submission to Chaos, a lead paragraph needs to be added here.}

@@

\section*{Acknowledgements}
Parts of this work have been financially supported by the National Natural Science Foundation of China (grant no. 11872182 and 11835003). YZ further acknowledges the support by Shanghai Municipal Science and Technology Major Project (grant no. 2018SHZDZX01) and ZJ Lab.
